### 2018-07-12 Qualifying Exam/conclusion.tex
so many changes to the dissertation, and Other!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter[Conclusion]{Conclusion and Future Work}
\label{chap-conc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    Inspired by large-scale human experiments with swarms of robots under global control,  this thesis investigated controllers that use only the mean and variance of a robot swarm. We proved that the mean position is controllable, and provided conditions under which variance is controllable.  We derived automatic controllers for each and a hysteresis-based switching control that controls the mean and variance of a robot swarm.  We employed these controllers as primitives for a block-pushing task. 
%These parts were done as my masters work, and has been published or submitted into two conference papers and is in draft in one IJRR journal paper. Fig.~\ref{fig:timeline} shows the timeline plan I have. The green parts in the timeline has already been done, and I will be working on the others during my PhD years.

@@

  \section{Future Work: Gathering Problem}
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction. These activities require robots that behave intelligently. Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire robot population. It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location. 
However, it is often possible to sense global properties of the group, such as mean position and variance. In our experiments, we showed that controlling mean is possible, and controlling variance with some constraints is possible. One of the constraints was that we had assumed that the swarm is clustered together and the swarm was trying to keep the variance small. If we did not assume that and we had a world like Fig.~\ref{fig:sshaped} or if during the experiment, a fraction of the swarm makes a new branch and put the swarm to two or more clusters by hitting the obstacles in their path, we could not gather them together again, and our algorithm for controlling variance was failed. So the next problem is to design an algorithm that will gather all the robots to an specific region with global input so that we can have the minimum variance. \\

@@

\section{Future Work: Torque Control}

Another remaining problem is that sometimes we do not want the block to rotate during the experiment, or we want to rotate it to some angle. The problem statement is how to control torque of the object, and discussing what features of the object make torque controllable. Fig.~\ref{fig:torque} shows an example of torque control.
    \begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{torque.png}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:torque} How can we control torque of the object with a swarm of robots and a shared input? }
\end{figure}
    % We should also control the covariance $\sigma_xy$ and higher moments of the distribution

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion and Future Work}\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    Inspired by large-scale human experiments with swarms of robots under global control,  this thesis investigated controllers that use only the mean and variance of a robot swarm. We proved that the mean position is controllable, and provided conditions under which variance is controllable.  We derived automatic controllers for each and a hysteresis-based switching control that controls the mean and variance of a robot swarm.  We employed these controllers as primitives for a block-pushing task. 
%These parts were done as my masters work, and has been published or submitted into two conference papers and is in draft in one IJRR journal paper. Fig.~\ref{fig:timeline} shows the timeline plan I have. The green parts in the timeline has already been done, and I will be working on the others during my PhD years.

@@

  \subsection{Gathering Problem}
Large populations of micro- and nanorobots are being produced in laboratories around the world, with diverse potential applications in drug delivery and construction. These activities require robots that behave intelligently. Limited computation and communication rules out autonomous operation or direct control over individual units; instead we must rely on global control signals broadcast to the entire robot population. It is not always practical to gather pose information on individual robots for feedback control; the robots might be difficult or impossible to sense individually due to their size and location. 
However, it is often possible to sense global properties of the group, such as mean position and variance. In our experiments, we showed that controlling mean is possible, and controlling variance with some constraints is possible. One of the constraints was that we had assumed that the swarm is clustered together and the swarm was trying to keep the variance small. If we did not assume that and we had a world like Fig. ~\ref{fig:sshaped} or if during the experiment, a fraction of the swarm makes a new branch and put the swarm to two or more clusters by hitting the obstacles in their path, we could not gather them together again, and our algorithm for controlling variance was failed. So the next problem is to design an algorithm that will gather all the robots to an specific region with global input so that we can have the minimum variance. \\

@@

\subsection{Torque Control}

Another problem that still remains, is that sometimes we don't want the block to rotate during the experiment, or we want to rotate it in some angles. The problem statement is how to control torque of the object, and what are the features of the object that torque is controllable? Fig.~\ref{fig:torque} shows an example of torque control.
    \begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{torque.png}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:torque} How can we control torque of the object with a swarm of robots and a shared input? }
\end{figure}
    % We should also control the covariance $\sigma_xy$ and higher moments of the distribution

### 2018-07-12 Qualifying Exam/intro.tex
so many changes to the dissertation, and Other!

\chapter[Introduction]{Introduction}
\label{chap-intro}

\section{Shaping a Swarm Using a Shared Control Input}\label{sec:Intro}
Micro- and nanorobotics can be manufactured in large numbers.
Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
 The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite incredible under-actuation.  
Additionally, it is not always possible to gather pose information on each robot for feedback control. 
Robots might be difficult or impossible to sense individually due to their size and location. 
For example, micro-robots are smaller than the minimum resolution of a clinical MRI-scanner~\cite{martel2014computer}.
However, it is often possible to sense global properties of the group, such as mean position and variance. 
To make progress in automatic control with global inputs, we present swarm manipulation controllers requiring only mean and variance measurements of the robot's positions.  These controllers are used as primitives to perform a block-pushing task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.
\begin{figure}
\centering
\begin{overpic}[width=1.0\columnwidth]{BlockBotView2.eps}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.  Here is a mockup of a swarm of hardware robots(kilobots) that pushes a green block toward the goal. %See video attachment~\cite{ShivaVideo2015}.
}
\end{figure}
A limitation was that variance control could only compress a swarm along the world $x$ and $y$ axes.  This means the swarm could not navigate workspaces with narrow corridors with other orientations, such as those shown in Fig.\ \ref{fig:covFriction}.
Challenges like these require a controller that regulates the swarm's position covariance, $\sigma_{xy}$. 
This thesis continues to prove  that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), implements these position control algorithms in simulation (Section \ref{sec:simulation}) and on a hardware setup with up to 64 robots (Section \ref{sec:expResults}), and ends with directions for future research.

\begin{figure}
\centering
\begin{overpic}[width=0.9\columnwidth]{Leaf2.jpg}\end{overpic}
\caption{\label{fig:vascularNetwork}%Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.  
Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging due to many obstacles. This thesis demonstrates how friction with walls can be used to change the shape of a swarm.} %TODO: save this as a pdf
\end{figure}

>>>>

\section{Introduction}\label{sec:Intro}
Micro- and nanorobotics can be manufactured in large numbers.
Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
 The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite incredible under-actuation.  
Additionally, it is not always possible to gather pose information on each robot for feedback control. 
Robots might be difficult or impossible to sense individually due to their size and location. 
For example, micro-robots are smaller than the minimum resolution of a clinical MRI-scanner~\cite{martel2014computer}.
However, it is often possible to sense global properties of the group, such as mean position and variance. 
To make progress in automatic control with global inputs, we present swarm manipulation controllers requiring only mean and variance measurements of the robot's positions.  These controllers are used as primitives to perform a block-pushing task illustrated in Fig.~\ref{fig:bigPictureMeanAndVarianceForSwarm}.
\begin{figure}
\centering
\begin{overpic}[width=1.0\columnwidth]{BlockBotView2.eps}\end{overpic}
%\todo{I like the 'target' symbol, but it is not self-documenting.  We need a legend explaining the min and max variance ellipses, the goal region, the variance, the mean, the object COM, and the target mean position.  I think these are easiest to make in powerpoint.
%Please use the same color and line style for the variance min and max as you use in Figure 4.
%}
%{blockpushingImageWithMeanAndVarianceOverlay.png}
\caption{\label{fig:bigPictureMeanAndVarianceForSwarm} A swarm of robots, all controlled by a uniform force field, can be effectively controlled by a hybrid controller that knows only the first and second moments of the robot distribution.  Here is a mockup of a swarm of hardware robots(kilobots) that pushes a green block toward the goal. See video attachment~\cite{ShivaVideo2015}.}
\end{figure}
A limitation was that variance control could only compress a swarm along the world $x$ and $y$ axes.  This means the swarm could not navigate workspaces with narrow corridors with other orientations, such as those shown in Fig.\ \ref{fig:covFriction}.
Challenges like these require a controller that regulates the swarm's position covariance, $\sigma_{xy}$. 
Because of this reason, this paper continues to prove  that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots (Section \ref{sec:PostionControl2Robots}), implements these position control algorithms in simulation (Section \ref{sec:simulation}) and on a hardware setup with up to 64 robots (Section \ref{sec:expResults}), and ends with directions for future research.

Micro- and nano-robots can be manufactured in large numbers.
Our vision is for large swarms of robots remotely guided 1) through the human body, to cure disease, heal tissue, and prevent infection and 2) ex vivo to assemble structures in parallel. 
 For each application, large numbers of micro robots are required  to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
 The biggest barrier to this vision is a lack of control techniques that can reliably exploit large populations despite incredible under-actuation.

\begin{figure}
\centering
\begin{overpic}[width=0.9\columnwidth]{Leaf2.jpg}\end{overpic}
\caption{\label{fig:vascularNetwork}Vascular networks are common in biology such as the circulatory system and cerebrospinal spaces, as well as in porous media including sponges and pumice stone.  Navigating a swarm using global inputs, where each member receives the same control inputs, is challenging
due to the many obstacles. This paper demonstrates how friction with walls can be used to change the shape of a swarm.} %TODO: save this as a pdf
\end{figure}

### 2018-07-12 Shiva Dissertation/chap/intro.tex
so many changes to the dissertation, and Other!

\begin{figure}
\centering
\begin{overpic}[width=0.5\columnwidth]{MainExpFig.pdf}\end{overpic}
\begin{overpic}[width=0.465\columnwidth]{CoverPhoto.pdf}
\end{overpic}\\

\begin{overpic}[width=0.45\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
\begin{overpic}[width=0.45\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
\vspace{3em}
\caption{\label{fig:chapters} 
Top left: Chapter 2 discusses steering a swarm of particles using global inputs and swarm statistics. Top right: Chapter 3 goes deep into the torque control. Bottom: Chapter 4 demonstrates algorithms to position two robots. %Bottom right: the magnetic setup that is used to manipulate magnetic particles when they both get the same control input.
}
\end{figure}

>>>>

\begin{figure}
\centering
\begin{overpic}[width=0.39\columnwidth]{MainExpFig.pdf}\end{overpic}
\begin{overpic}[width=0.365\columnwidth]{CoverPhoto.pdf}
\end{overpic}\\

\begin{overpic}[width=0.35\columnwidth]{firstpicLeft.pdf}\put(28,-10){workspace}\end{overpic}
\begin{overpic}[width=0.35\columnwidth]{magneticsetup.pdf}\put(22,-8){magnetic setup}\end{overpic}
\vspace{3em}
\caption{\label{fig:chapters} 
Top left: Chapter 2 discusses steering a swarm of particles using global inputs and swarm statistics. Top right: Chapter 3 goes deep into the torque control. Bottom: Chapter 4 demonstrates algorithms to position two robots. %Bottom right: the magnetic setup that is used to manipulate magnetic particles when they both get the same control input.
}
\end{figure}

### 2018-07-12 Shiva Dissertation/chap2/experiment.tex
so many changes to the dissertation, and Other!

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}  \left[ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u}) \right].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_i) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} \left[r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})\right] \label{eq:ValueIteration} \nonumber\\
\text{end}&  %page 502
\end{align}
Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25. 
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object, as shown in Fig.~\ref{fig:potentialField}. The approach is similar to \cite{spong2008robot}, chapter 5.
The repulsive potential field is centered at $\mathbf{b}$, the object's COM, and is active in a circular sector of angular width 2$\theta$ and radius $d_0$ aligned with $\mathbf{D}(\mathbf{b})$. $\mathbf{D}(\mathbf{b})$ is the desired direction of motion from~\eqref{eq:OptimalPolicy}.
\begin{align}
\mathbf{d} &=  [\bar{x},\bar{y}] - \mathbf{b}, \qquad   \phi =\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot  \mathbf{d}  }{\norm{\mathbf{D}(\mathbf{b})} \cdot \norm{\mathbf{d}} } \right), \nonumber \\ 
F_{\rm{att}} &= -\zeta \frac{\mathbf{d}  }{ \norm{\mathbf{d}}}, \nonumber\\
 F_{\rm{rep}} &=  \left\{
\begin{array}{ll}
      \eta( \frac{1}{\norm{\mathbf{d}}}- \frac{1}{d_0}) \frac{1}{\norm{\mathbf{d}}^2} \mathbf{d}, ~& \norm{\mathbf{d}} \leq d_0~\wedge~\phi <  \theta \\
      0, & \textrm{ otherwise} \nonumber \\
\end{array},
\right.\\
F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}, \nonumber\\
[x_{\rm{goal}}&, y_{\rm{goal}}] =  [\bar{x},\bar{y}]  + \frac{F_{\rm{pot}}}{\norm{F_{\rm{pot}}}}.
\end{align}
Here  $\eta$ and $\zeta$ are positive parameters that scale the forces and $\norm{\mathbf{d}}$ is the distance from the swarm mean $ [\bar{x},\bar{y}]$ to the object COM.
In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $d_0 = 3$. Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $d_0 = 7.5$.

>>>>

% A corresponding reward function gives a high reward to the goal state,  
% a large negative reward to states including obstacles, and a small negative reward to all other states.
The reward function $r(x,\mathbf{u})$ is defined as
\begin{align}
r(x,\mathbf{u}) &=  \left\{
\begin{array}{ll}
     +100, &  \textrm{if } \mathbf{u} \textrm{ leads to goal state,}\\
      -100, & \textrm{if } \mathbf{u} \textrm{ leads to an obstacle state,} \\
      -1, & \textrm{otherwise}\\
\end{array} 
\right.
\end{align}
 where $x$ is the current state and $\mathbf{u}$ is the action.   %cumulative expected
  Value iteration computes $\hat{V}(x)$, the expected discounted sum reward if the optimal policy is implemented, for the object starting in each state $x$. The optimal policy is %the control action that, in probability, results in the largest value:
   \begin{align} \mathbf{D}(x) = \arg\max_{\mathbf{u}}  \left[ r(x,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x, \mathbf{u}) \right].  \label{eq:OptimalPolicy}
   \end{align}
   The value function $\hat{V}(x_i) $ is calculated by computing the value $\hat{V}$ for all $N$ states and iterating until convergence:
\begin{align}
\text{for }&\text{$i=1$ to $N$ do} \nonumber \\
&\hat{V} (x_i) = \gamma \max_{\mathbf{u}} \left[r(x_i,\mathbf{u}) + \sum\limits_{j=1}^N \hat{V}(x_j) p(x_j | x_i,\mathbf{u})\right] \label{eq:ValueIteration} \nonumber\\
\text{end}& \textrm{.}  %page 502
\end{align}
Our probabilistic motion model $p(x_j | x, \mathbf{u})$ assumed the object moved in the commanded direction $\mathbf{u}$ half of the time but $+45^\circ$ with probability 0.25 and $-45^\circ$ with probability 0.25. 
In our experiments $\gamma = 0.97$, and \eqref{eq:ValueIteration} was iterated 200 times. A {\sc Matlab} implementation is available at \cite{Becker2015MDP}.
%Rather than a shortest distance to goal, $\nabla \mathbf{M}_{MDP}$  gives the approximate minimum cost to goal.

@@

When the swarm is in front of the object, control law \eqref{eq:PDcontrolPosition} pushes the object backwards.  To fix this, we implement a potential field approach that attracts the swarm to the intermediate goal, but repulses the swarm from in front of the object, as shown in Fig.~\ref{fig:potentialField}. The approach is similar to \cite{spong2008robot}, chapter 5.
The repulsive potential field is centered at $\mathbf{b}$, the object's COM, and is active in a circular sector of angular width 2$\theta$ and radius $d_0$ aligned with $\mathbf{D}(\mathbf{b})$. $\mathbf{D}(\mathbf{b})$ is the desired direction of motion from~\eqref{eq:OptimalPolicy}. Therefore, we compute the goal locations with the following equations:
\begin{align}
\mathbf{d} &=  [\bar{x},\bar{y}] - \mathbf{b}, \qquad   \phi =\cos^{-1}\left( \frac{ \mathbf{D}(\mathbf{b}) \cdot  \mathbf{d}  }{\norm{\mathbf{D}(\mathbf{b})} \cdot \norm{\mathbf{d}} } \right), \nonumber \\ 
F_{\rm{att}} &= -\zeta \frac{\mathbf{d}  }{ \norm{\mathbf{d}}}, \nonumber\\
 F_{\rm{rep}} &=  \left\{
\begin{array}{ll}
      \eta( \frac{1}{\norm{\mathbf{d}}}- \frac{1}{d_0}) \frac{1}{\norm{\mathbf{d}}^2} \mathbf{d}, ~& \norm{\mathbf{d}} \leq d_0~\wedge~\phi <  \theta \\
      0, & \textrm{ otherwise} \nonumber \\
\end{array},
\right.\\
F_{\rm{pot}} &= F_{\rm{att}} + F_{\rm{rep}} \label{eq:potentialfield}, \nonumber\\
[x_{\rm{goal}}&, y_{\rm{goal}}] =  [\bar{x},\bar{y}]  + \frac{F_{\rm{pot}}}{\norm{F_{\rm{pot}}}}.
\end{align}
Here  $\eta$ and $\zeta$ are positive parameters that scale the forces and $\norm{\mathbf{d}}$ is the distance from the swarm mean $ [\bar{x},\bar{y}]$ to the object COM.
In simulations, $\theta =  \pi/2$,  $\eta  = 75$, $\zeta = 2$ and $d_0 = 3$. Because the kilobots have a slower time constant, they use $\theta =  \pi/2$,  $\eta  = 50$, $\zeta = 1$ and $d_0 = 7.5$.

### 2018-07-12 Shiva Dissertation/chap2/realExperiment.tex
so many changes to the dissertation, and Other!

\subsection{Environmental setup}  
In these experiments as shown in Fig.~\ref{fig:setup}, we used $n$=100 kilobots and a 1.5 m$\times$1.2 m whiteboard as the workspace. 
LED floodlights were placed 1.5 m above the table on the sides and corners of a square with 6 m sides. An Arduino Uno connected to an 8-relay shield controlled the lights.

@@

\begin{figure*}
%\renewcommand{\figwid}{5.25cm}
\begin{center}
	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
	\begin{overpic}[width=0.7\columnwidth]{SetUp.pdf}%\put(1,75){A}
	\end{overpic}\\
	\vspace{3em}
	\begin{overpic}[width=0.33\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
	\end{overpic}
	\begin{overpic}[width=0.6\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
	\end{overpic}
\end{center}
\caption{\label{fig:setup}
Hardware platform. At right are the shapes used for hardware experiments and a visualization of the potential field. }
\end{figure*}

@@

\paragraph{Swarm mean control (hardware experiment)}

Unlike the PD controller \eqref{eq:PDcontrolPosition}, we cannot command a force input to the kilobots.  Instead, control is given by turning on one of eight lights.  The kilobots run a phototaxis routine where they search for an orientation that aligns them with the light source, and then move with an approximately constant velocity toward this light.   The kilobots oscillate along this orientation because they only have one light detector.

>>>>

\subsection{Environmental setup}  
In these experiments as shown in Fig.~\ref{fig:setup2}, we used $n$=100 kilobots and a 1.5 m$\times$1.2 m whiteboard as the workspace. 
LED floodlights were placed 1.5 m above the table on the sides and corners of a square with 6 m sides. An Arduino Uno connected to an 8-relay shield controlled the lights.

@@

\begin{figure*}
%\renewcommand{\figwid}{5.25cm}
\begin{center}
	%\includegraphics[width=0.5\columnwidth]{SetUp.pdf}
	\begin{overpic}[width=0.7\columnwidth]{SetUp.pdf}%\put(1,75){A}
	\end{overpic}\\
	\vspace{3em}
	\begin{overpic}[width=0.33\columnwidth]{AllShapesR.pdf}%\put(-0,85){B}
	\end{overpic}
	\begin{overpic}[width=0.6\columnwidth]{PotentialExp.pdf}%\put(-0,85){C}
	\end{overpic}
\end{center}
\caption{\label{fig:setup2}
Hardware platform. At right are the shapes used for hardware experiments and a visualization of the potential field. }
\end{figure*}

@@

\subsection{Swarm mean control (hardware experiment)}

Unlike the PD controller \eqref{eq:PDcontrolPosition}, we cannot command a force input to the kilobots.  Instead, control is given by turning on one of eight lights.  The kilobots run a phototaxis routine where they search for an orientation that aligns them with the light source, and then move with an approximately constant velocity toward this light. % The kilobots oscillate along this orientation because they only have one light detector.

### 2018-07-12 Shiva Dissertation/chap2/theory.tex
Shiva PhD dissertation
Still requires some edits.
dissertation as I sent to committee members.
so many changes to the dissertation, and Other!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Global Control Laws for a holonomic swarm}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We represent particles as holonomic robots that move in the 2D plane. We want to control position and velocity of the particles. 
First, assume a noiseless system containing one particle with mass $m$.
 Our inputs are global forces $[u_x,u_y]^\top$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
The state-space representation in standard form is: 
\begin{align}\label{eq:stdform}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t).
\end{align}

@@

Controlling the variance requires being able to increase and decrease the variance.  We will list a sufficient condition for each. 
Microscale systems are affected by unmodelled dynamics. 
These unmodelled dynamics are dominated by Brownian noise, as described in~\cite{einstein1956investigations}. 
To model this~\eqref{eq:stdform} must be modified as follows:
\begin{align}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t),
\end{align}
where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$. Given a large obstacle-free workspace with $\mathbf{u}(t)= 0$, a \emph{Brownian noise} process increases the variance linearly with time.
\begin{equation}
\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t))  = W \bm{\varepsilon},
\quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t.
\end{equation}
 If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board as in~\cite{Becker2013b}.

@@

Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}),   \label{eq:PDcontrolVariance}\\
u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the particle at $x=0$, and chooses a reference goal position such that the swarm, if uniformly distributed between 0 and $\ell$, would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.  A similar argument applies to~\eqref{eq:PDcontrolVarianceIny}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Global control laws for a holonomic swarm}
\label{sec:theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We represent particles as holonomic robots (robots that can move freely in any direction) that move in the 2D plane.
 We want to control position and velocity of the particles. 
First, assume a noiseless system containing one particle with mass $m$.
 Our inputs are global forces $[u_x,u_y]^\top$. We define our state vector $\mathbf{x}(t)$ as the $x$ position, $x$ velocity, $y$ position and $y$ velocity.
The state-space representation in standard form is: 
\begin{align}\label{eq:stdform}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t),
\end{align}

@@

Controlling the variance requires being able to increase and decrease the variance.  We will list a sufficient condition for each. 
Microscale systems are affected by unmodelled dynamics. 
These unmodelled dynamics are dominated by Brownian noise, as described in~\cite{einstein1956investigations}. 
To model this~\eqref{eq:stdform} must be modified as follows:
\begin{align}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t),
\end{align}
where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$. Given a large obstacle-free workspace with $\mathbf{u}(t)= 0$, a \emph{Brownian noise} process increases the variance linearly with time and can be calculated by
\begin{equation}
\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t))  = W \bm{\varepsilon},
\quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t.
\end{equation}
 If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board as in~\cite{Becker2013b}.

@@

Similar to the PD controller in Eq.~\eqref{eq:PDcontrolPosition} that controls the mean particle position, a controller to regulate the variance to $\sigma^2_{\rm{ref}}$ is:
\begin{align}
u_x &= K_{p}(x_{\rm{goal}}(\sigma^2_{\rm{ref}}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{x}) \textrm{       and}  \label{eq:PDcontrolVariance}\\
u_y &= K_{p}(y_{\rm{goal}}(\sigma^2_{\rm{ref}})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{\rm{ref}}-\sigma^2_{y}).  \label{eq:PDcontrolVarianceIny}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time. This controller requires a vertical and a horizontal wall.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the particle at $x=0$, and chooses a reference goal position such that the swarm, if uniformly distributed between 0 and $\ell$, would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{\rm{goal}}(\sigma^2_{\rm{ref}}) = \ell/2 = r + \sqrt{3\sigma^2_{\rm{ref}}}.
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{\rm{goal}}$ is translated.  A similar argument applies to~\eqref{eq:PDcontrolVarianceIny}.

### 2018-07-12 Shiva Dissertation/chap3/simulation.tex
so many changes to the dissertation, and Other!

\paragraph{Orientation of the object}
These simulations used a uniform density rectangle as the object. This object was 30$\times$ larger than the robots.
Using the pure torque control discussed in the previous paragraph, the orientation of the object can be controlled by applying force. 
The rectangular object is not pivoted, so it moves in addition to rotating. 
 The swarm still may split into multiple components.
  We use the hysteresis variance control from \cite{ShahrokhiIROS2015}  to gather the swarm when its variance grows too large. 
  The following control law chooses a goal position to regulate the orientation of the object.

@@

Fig. \ref{fig:OrientCont} illustrates this controller with different starting positions. When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace. Code is available at \cite{Shahrokhi16Orient}.

>>>>

\paragraph{Orientation of the object}
These simulations used a uniform density rectangle as the object. This object was 30$\times$ larger than the robots.
Using the pure torque control discussed in the previous paragraph, the orientation of the object can be controlled by applying force. 
The rectangular object is not pivoted, so it moves in addition to rotating. 
 The swarm still may split into multiple components.
  We use the hysteresis variance control from \cite{ShahrokhiIROS2015}  to gather the swarm when its variance grows too large. 
  The control law that chooses a goal position to regulate the orientation of the object is

@@

Fig. \ref{fig:OrientCont} illustrates this controller with different starting positions. When the plot traces are constant the swarm is no longer pushing the object and instead is being regathered in a corner of the workspace.% Code is available at \cite{Shahrokhi16Orient}.
